Randomly-Scoped Lambda Calculus

### 2013-01-26 RSLC/paper/paper.tex
Finish RSLC paper, I think

\section{Language Definition}

@@

The value $N = 19$ yields the $\mathsf{d}20$ combinator, which generates the random distribution shown in Figure~\ref{fig:d20}. Prior work has shown this combinator to be quite useful in tabletop roleplaying games~\cite{dnd-mm,dnd-phb,dnd-dmg}.

@@

@@

\newpage

@@

\section{Discussion of Substitution Rules}
\begin{figure*}[t]
	\fbox{$\subst{e_0}{x}{e}$}
	\TypeRules{
		\SubstLamCapture \qquad \SubstLam \qquad
		\SubstApp \\
		\SubstVarX \qquad \SubstVarY \\
		\SubstCaseCapture \\ \SubstCase \\
		\SubstZero \qquad \SubstSuc \\
		\SubstLetCapture \qquad \SubstLet
	}
	\fbox{$\capture{e_0}{x}{e}$}
	\TypeRules{
		\CaptureLam \qquad \CaptureApp \qquad
		\CaptureVarX \qquad \CaptureVarY \\
		\CaptureCase \\
		\CaptureZero \qquad \CaptureSuc \qquad \CaptureLet
	}
	\caption{Complete RSLC substitution and capture rules.}
	\label{fig:subst-complete}
\end{figure*}

>>>>

\section{Language Definition}
\label{sec:definition}

@@

The value $N = 19$ yields the $\mathsf{d}20$ combinator, which generates the random distribution shown in Figure~\ref{fig:d20} above. Prior work has shown this combinator to be quite useful in tabletop roleplaying games~\cite{dnd-mm,dnd-phb,dnd-dmg}.

@@

\section*{Acknowledgement}

Michael Sullivan is, of course, the ``Sully'' that I mentioned (and will in the appendix continue to mention).

@@

\onecolumn

@@

\section{Discussion of Substitution Rules}

Section~\ref{sec:definition} shows some ``most important'' example substitution rules, saving the complete set of rules for the appendix, which is where we are now, hence Figure~\ref{fig:subst-complete} below.

Now, Sully thought my two-judgement scheme was gross, and proposed his own, which I think is operationally equivalent.
Instead of switching to a different judgement the first time you traverse a matching binder (and annotating variables with expression-lists), in you always carry around a counter, which you increment when crossing matching binders and which represents a ``substitution probability'' (really the inverse thereof).

Figure~\ref{fig:sully} shows rules for this alternate substitution scheme.
In these rules, the notation $\sully{e_0}{x}{p}{e}$ represents substituting with probability $\sullycolor{1/p}$, and the default substitution used by the evaluation rules, $\subst{e_0}{x}{e}$, is equivalent to substituting with probability $1$, i.e., $\sully{e_0}{x}{1}{e}$.
Also, of course, variables don't have lists attached; they are $x$ rather than $x_{\rep{e}}$.
While this alternate system eliminates the clunkiness of annotating variables, it has its own clunkiness because the rules \textsc{Subst2-Var-X-Yes} and \textsc{Subst2-Var-X-No} must have explicit probabilities for which of the two of them will apply.

\begin{figure*}[h!]
	\fbox{$\subst{e_0}{x}{e}$}
	\TypeRules{
		\SubstLamCapture \qquad \SubstLam \qquad
		\SubstApp \\
		\SubstVarX \qquad \SubstVarY \\
		\SubstCaseCapture \\ \SubstCase \\
		\SubstZero \qquad \SubstSuc \\
		\SubstLetCapture \qquad \SubstLet
	}
	\fbox{$\capture{e_0}{x}{e}$}
	\TypeRules{
		\CaptureLam \qquad \CaptureApp \qquad
		\CaptureVarX \qquad \CaptureVarY \\
		\CaptureCase \\
		\CaptureZero \qquad \CaptureSuc \qquad \CaptureLet
	}
	\caption{Complete RSLC substitution and capture rules.}
	\label{fig:subst-complete}
\end{figure*}

\begin{figure*}[t]
	\fbox{$\sully{e_0}{x}{p}{e}$}
	\TypeRules{
		\SullyLamCapture \qquad \SullyLam \qquad \SullyApp \\
		\SullyVarXA \qquad \SullyVarXB \qquad \SullyVarY \\
		\SullyCaseCapture \\ \SullyCase \\
		\SullyZero \qquad \SullySuc \\
		\SullyLetCapture \qquad \SullyLet
	}
	\caption{Alternate substitution rules for RSLC. These maintain a counter for the number of binders traversed and uses explicit probabilities for randomness rather than selecting uniformly from a list.}
	\label{fig:sully}
\end{figure*}
